---
title: "Figure_1"
output: html_document
date: "2024-01-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This script inputs six rds files (3 injured, 3 uninjured) of lumbar- projecting neurons.
#It performs QC, combines and normalizes the files, then assigns and identity to clusters based on marker genes.

```{r}
library('Seurat')
library('ggplot2')
library('plyr')
library('dplyr')
```
#Load the Input data 
```{r}
Sample1.data <- Read10X(data.dir = "1_SourceFiles/LU2/")
Sample1 <- CreateSeuratObject(counts = Sample1.data, project = "Sample1", min.cells = 3, min.features = 1000)
Sample1 

VlnPlot(Sample1, features = c("nFeature_RNA"))
Sample1 <- subset(Sample1, subset = nFeature_RNA > 1000 & nFeature_RNA < 6000)
Sample1 

Sample2.data <- Read10X(data.dir = "1_SourceFiles/LU3")
Sample2 <- CreateSeuratObject(counts = Sample2.data, project = "Sample2",min.cells = 3, min.features = 1000)
Sample2 

VlnPlot(Sample2, features = c("nFeature_RNA"))
Sample2 <- subset(Sample2, subset = nFeature_RNA > 1000 & nFeature_RNA < 8000)
Sample2 

Sample3.data <- Read10X(data.dir = "1_SourceFiles/LU4")
Sample3 <- CreateSeuratObject(counts = Sample3.data, project = "Sample3", min.cells = 3, min.features = 1000)
Sample3

VlnPlot(Sample3, features = c("nFeature_RNA"))
Sample3 <- subset(Sample3, subset = nFeature_RNA > 1000 & nFeature_RNA < 7000)
Sample3

Sample4.data <- Read10X(data.dir = "1_SourceFiles/LI1")
Sample4 <- CreateSeuratObject(counts = Sample4.data, project = "Sample4",min.cells = 3, min.features = 1000)
Sample4 
VlnPlot(Sample4, features = c("nFeature_RNA"))
Sample4 <- subset(Sample4, subset = nFeature_RNA > 1000 & nFeature_RNA < 5500)
Sample4 

Sample5.data <- Read10X(data.dir = "1_SourceFiles/LI2")
Sample5 <- CreateSeuratObject(counts = Sample5.data, project = "Sample5", min.cells = 3, min.features = 1000)
Sample5
VlnPlot(Sample5, features = c("nFeature_RNA"))
Sample5 <- subset(Sample5, subset = nFeature_RNA > 1000 & nFeature_RNA < 7500)
Sample5

Sample6.data <- Read10X(data.dir = "1_SourceFiles/LI3")
Sample6 <- CreateSeuratObject(counts = Sample6.data, project = "Sample6", min.cells = 3, min.features = 1000)
Sample6 
VlnPlot(Sample6, features = c("nFeature_RNA"))
Sample6 <- subset(Sample6, subset = nFeature_RNA > 1000 & nFeature_RNA < 7500)
Sample6 
```

#Merge all the data set

```{r}
LMN_merge.list <- list (Sample1, Sample2, Sample3, Sample4, Sample5, Sample6)

LMN_merge.list <- lapply(X=LMN_merge.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

features <- SelectIntegrationFeatures(object.list = LMN_merge.list)

LMN_merge.anchors <- FindIntegrationAnchors(object.list = LMN_merge.list,
                                            anchor.features = features)

LMN <- IntegrateData(anchorset = LMN_merge.anchors)

DefaultAssay(LMN) <- "integrated"

LMN <- ScaleData(LMN, verbose = FALSE)

LMN <- RunPCA(LMN, npcs = 50, verbose = FALSE)

LMN <- RunUMAP(LMN, reduction = "pca", dims = 1:30)

LMN <- FindNeighbors(LMN, reduction = "pca", dims = 1:30)

LMN <- FindClusters(LMN, resolution = 0.5)

DefaultAssay(LMN) <- "RNA"

LMN

DimPlot(LMN, group.by = 'seurat_clusters', split.by = 'orig.ident',
        label = TRUE, label.size = 5) + NoLegend()
```

#Remove the clearly stressed, dying, or artificially low cell count clusters.

```{r}
FeaturePlot(LMN, c("Ddit3"), label = TRUE, 
            label.size = 5, repel = TRUE, order = TRUE, split.by = "orig.ident")

FeaturePlot(LMN, c("nFeature_RNA"), label = TRUE, 
            label.size = 5, repel = TRUE, order = TRUE) 

VlnPlot(LMN, "nFeature_RNA", pt.size = 0) + NoLegend() +
  theme(axis.text.x = element_text(angle = 0, hjust = .5)) +
  theme(text = element_text(size = 20))

VlnPlot(LMN, "Ddit3", pt.size = 0) + NoLegend() +
  theme(axis.text.x = element_text(angle = 0, hjust = .5)) +
  theme(text = element_text(size = 20))
```
#Based on the above QC metrics, cluster 4 and 11 are removed for displaying high death even in uninjured samples, and clusters 12, 24, and 25 for low feature counts
```{r}
Idents(LMN) <- "seurat_clusters"

LMN1 <- subset(x = LMN, idents = c("4", "11", "12", "24","25"), invert = TRUE)

#A confidence check visualization to confirm proper removal

DimPlot(LMN1, group.by = 'seurat_clusters', split.by = 'orig.ident', label = TRUE, label.size = 5) + NoLegend()
```
```{r}
#Re normalize the new dataset that lacks the dying and low-count clusters cells

DefaultAssay(LMN1) <- "RNA"

LMN1 <- NormalizeData(LMN1, normalization.method = "LogNormalize", scale.factor = 10000)

LMN1 <- FindVariableFeatures(LMN1, selection.method = "vst", nfeatures = 2000)

all.genes <- rownames(LMN1)

LMN1 <- ScaleData(LMN1, features = all.genes)

LMN1 <- RunPCA(LMN1, features = VariableFeatures(object = LMN1), npcs = 50)

LMN1 <- FindNeighbors(LMN1, dims = 1:30)
LMN1 <- FindClusters(LMN1, resolution = 0.5)

LMN1 <- RunUMAP(LMN1, dims = 1:30)

LMN1 

```

```{r}
FeaturePlot(LMN1, ("Gad2"), 
            label = TRUE, label.size = 3,
            repel = FALSE, order = TRUE, )

DimPlot(LMN1, group.by = 'seurat_clusters',  label = TRUE, label.size = 5) + NoLegend()
```

```{r}
DotPlot(LMN1, features = c("Mef2c", "Camk2a", "Ptk2b", "Satb2", 
                           "Crym", "Fezf2", "Bcl6", "Slc30a3", "Sulf1"),
        dot.scale = 4, dot.min = 0) + coord_flip() +
  theme(axis.text.x = element_text(angle = 0, hjust = .5),
        axis.text = element_text(size = 10))
```

```{r}
VlnPlot(LMN1, "Satb2", pt.size = 0) + NoLegend() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 15))
```
The conclusion is that clusters 0,1, and 3 are CST but 15 and 18 are intracortical 
non-CST contamination that should be removed
Remove clusters 15 and 18, creating LMN_Final
```{r}
Idents(LMN1) <- "seurat_clusters"
levels(LMN1)

LMN_Final <- subset(LMN1, idents = c("15", "18"), invert = TRUE)
```

```{r}
#Renormalize and cluster LMN_Final now that cluster 15 and 18 are removed (probably won't change much)

DefaultAssay(LMN_Final) <- "RNA"

LMN_Final <- NormalizeData(LMN_Final, normalization.method = "LogNormalize", scale.factor = 10000)

LMN_Final <- FindVariableFeatures(LMN_Final, selection.method = "vst", nfeatures = 2000)

all.genes <- rownames(LMN_Final)

LMN_Final <- ScaleData(LMN_Final, features = all.genes)

LMN_Final <- RunPCA(LMN_Final, features = VariableFeatures(object = LMN_Final), npcs = 50)

LMN_Final <- FindNeighbors(LMN_Final, dims = 1:30)
LMN_Final <- FindClusters(LMN_Final, resolution = 0.5)

LMN_Final <- RunUMAP(LMN_Final, dims = 1:30)

LMN_Final

DimPlot(LMN_Final, group.by = 'seurat_clusters', split.by = 'orig.ident',
        label = TRUE, label.size = 3) + NoLegend()

DimPlot(LMN_Final, group.by = 'seurat_clusters',
        label = TRUE, label.size = 5) + NoLegend()

VlnPlot(LMN_Final, "Rmst", pt.size = 0) + NoLegend() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 15)) #Sized for display

DotPlot(LMN_Final, features = c("Mef2c", "Camk2a", "Ptk2b", "Satb2", 
                           "Crym", "Fezf2", "Bcl6", "Slc30a3", "Sulf1"),
        dot.scale = 4, dot.min = 0) + coord_flip() +
  theme(axis.text.x = element_text(angle = 0, hjust = .5),
        axis.text = element_text(size = 10)) #Sized for display

FeaturePlot(LMN_Final, ("Camk2a"), 
            label = TRUE, label.size = 3,
            repel = FALSE, order = TRUE, )
```


```{r}
LMN_Final_DotPlot1 <- read.csv("2_Input_Lists/LMN_Final_DotPlot1.csv",sep = ",")

Idents(LMN_Final)  <- "seurat_clusters"

DefaultAssay(LMN_Final) <- "RNA"

DotPlot(LMN_Final, features = LMN_Final_DotPlot1$Marker, dot.scale = 9, dot.min = 0) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0, hjust = .5),
        axis.text = element_text(size = 10)) #Sized for display


FeaturePlot(LMN_Final, c("Sox14"), 
            label = TRUE, label.size = 3,
            repel = FALSE, order = TRUE, ) #500 x 1000
```

```{r}
#LMN_Final_1 <- JoinLayers(LMN_Final)
#LMN_Final_17 <- FindMarkers(LMN_Final_1,ident.1 = c("17"))
#LMN_Final_1=LMN_Final
#write.csv(x = LMN_Final_17, file = "LMN_Final_17.csv",     quote = FALSE)
```



```{r}
FeaturePlot(LMN_Final, c("Sox14"), split.by = 'orig.ident', 
            label = TRUE, label.size = 4,
            repel = FALSE, order = TRUE, ) #500 x 1000

VlnPlot(LMN_Final, "Ebf2", pt.size = 0) + NoLegend() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 15)) #Sized for display



#Map seurat_clusters to manual.cluster.1 (anatomical assignments)

LMN_Final_Map1 <- read.csv("2_Input_Lists/LMN_Final_Map1_manual_clusters.csv",
                           sep = ",")

Idents(LMN_Final) <- "seurat_clusters"

levels(LMN_Final)

LMN_Final@meta.data$manual.clusters.1 <- plyr::mapvalues(
  x = LMN_Final@meta.data$seurat_clusters,
  from = LMN_Final_Map1$seurat_clusters,
  to = LMN_Final_Map1$manual.cluster.1)

Idents(LMN_Final) <- "manual.clusters.1"

levels(LMN_Final)


#Order Seurat Clusters in a rostral-caudal sequence

LMN_Final@meta.data$seurat_clusters <- factor(
  x = LMN_Final@meta.data$seurat_clusters,
  levels = LMN_Final_Map1$seurat_clusters)

Idents(LMN_Final) <- "seurat_clusters"
levels(LMN_Final)


#Order manual.clusters.1 in a rostral-caudal sequence

LMN_Final_Order1 <- read.csv("2_Input_Lists/LMN_Final_Order1.csv",
                             sep = ",")

LMN_Final@meta.data$manual.clusters.1 <- factor(
  x = LMN_Final@meta.data$manual.clusters.1,
  levels = LMN_Final_Order1$Order)

Idents(LMN_Final) <- "manual.clusters.1"
levels(LMN_Final)


#Check identities with Dotplots

LMN_Final_DotPlot2 <- read.csv("2_Input_Lists/LMN_Final_Dotplot2.csv", sep = ",")


DefaultAssay(LMN_Final) <- "RNA"
```

```{r}
#DotPlot(LMN_Final, features = LMN_Final_DotPlot1$Marker, dot.scale = 4, dot.min = 0) +
#  coord_flip() +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1),
#        axis.text = element_text(size = 10)) #Sized for display

#DotPlot(LMN_Final, features = LMN_Final_DotPlot2$Marker, dot.scale = 4, dot.min = 0) +
#  coord_flip() +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 10)) #Sized for display


#Add information about injury status to LMN_Final

LMN_Final_Map2 <- read.csv("2_Input_Lists/LMN_Final_Map2.csv",
                           sep = ",")

Idents(LMN_Final) <- "orig.ident"
levels(LMN_Final)

LMN_Final@meta.data$injury.status <- plyr::mapvalues(
  x = LMN_Final@meta.data$orig.ident,
  from = LMN_Final_Map2$orig.ident,
  to = LMN_Final_Map2$injury.status)

Idents(LMN_Final) <- "injury.status"
levels(LMN_Final)

#Create cortical versus subcortical category
#For convenience I added these to the Map1 csv file as a new column

LMN_Final_Map1 <- read.csv("2_Input_Lists/LMN_Final_Map1_manual_clusters.csv",
                           sep = ",")

Idents(LMN_Final) <- "seurat_clusters"
levels(LMN_Final)

LMN_Final@meta.data$Cort_v_Subcort <- plyr::mapvalues(
  x = LMN_Final@meta.data$seurat_clusters,
  from = LMN_Final_Map1$seurat_clusters,
  to = LMN_Final_Map1$Cort_v_Subcort)

Idents(LMN_Final) <- "Cort_v_Subcort"
levels(LMN_Final)

#Create a manual.clusters.2, which provides larger groups, for example
#lateral and paraventricular hypothalamus are now just "HYP" and all the
#various brainstem clusters just become "MED" for medulla.

Idents(LMN_Final) <- "seurat_clusters"
levels(LMN_Final)

LMN_Final@meta.data$manual.clusters.2 <- plyr::mapvalues(
  x = LMN_Final@meta.data$seurat_clusters,
  from = LMN_Final_Map1$seurat_clusters,
  to = LMN_Final_Map1$manual.cluster.2)

Idents(LMN_Final) <- "manual.clusters.2"
levels(LMN_Final)

#Order manual.clusters.2

LMN_Final_Order2 <- read.csv("2_Input_Lists/LMN_Final_Order2.csv",
                             sep = ",")

LMN_Final@meta.data$manual.clusters.2 <- factor(
  x = LMN_Final@meta.data$manual.clusters.2,
  levels = LMN_Final_Order2$Order)

Idents(LMN_Final) <- "manual.clusters.2"
levels(LMN_Final)
```

```{r}
Idents(LMN_Final) <- "injury.status"
levels(LMN_Final)

Idents(LMN_Final) <- "manual.clusters.2"
levels(LMN_Final)

FeaturePlot(LMN_Final, c("Sox14"), split.by = 'injury.status', 
            label = TRUE, label.size = 4,
            repel = FALSE, order = TRUE, )

```

#DEG Analysis
```{r}
LMN_final_1=LMN_Final
LMN_Final <- JoinLayers(LMN_Final)

LMN_CST_InjVUninj <- FindMarkers(LMN_Final,
                                 subset.ident = "CST",
                                 group.by = "injury.status",
                                 ident.1 = "injured",
                                 ident.2 = "uninjured",
                                 logfc.threshold = 0,
                                 min.pct = 0.01)


write.csv(x = LMN_CST_InjVUninj, 
          file = "LMN_CST_InjvUninj.csv",
          quote = FALSE)


LMN_HYP_InjVUninj <- FindMarkers(LMN_Final,
                                 subset.ident = "HYP",
                                 group.by = "injury.status",
                                 ident.1 = "injured",
                                 ident.2 = "uninjured",
                                 logfc.threshold = 0,
                                 min.pct = 0.01)


write.csv(x = LMN_HYP_InjVUninj, 
          file = "LMN_HYP_InjvUninj.csv",
          quote = FALSE)

LMN_RN_InjVUninj <- FindMarkers(LMN_Final,
                                 subset.ident = "RN",
                                 group.by = "injury.status",
                                 ident.1 = "injured",
                                 ident.2 = "uninjured",
                                 logfc.threshold = 0,
                                 min.pct = 0.01)


write.csv(x = LMN_RN_InjVUninj, 
          file = "LMN_RN_InjvUninj.csv",
          quote = FALSE)


LMN_DP_InjVUninj <- FindMarkers(LMN_Final,
                                subset.ident = "DP",
                                group.by = "injury.status",
                                ident.1 = "injured",
                                ident.2 = "uninjured",
                                logfc.threshold = 0,
                                min.pct = 0.01)


write.csv(x = LMN_DP_InjVUninj, 
          file = "LMN_DP_InjvUninj.csv",
          quote = FALSE)

LMN_MED_InjVUninj <- FindMarkers(LMN_Final,
                                subset.ident = "MED",
                                group.by = "injury.status",
                                ident.1 = "injured",
                                ident.2 = "uninjured",
                                logfc.threshold = 0,
                                min.pct = 0.01)


write.csv(x = LMN_MED_InjVUninj, 
          file = "LMN_MED_InjvUninj.csv",
          quote = FALSE)

LMN_ALL_InjVUninj <- FindMarkers(LMN_Final,
                                 group.by = "injury.status",
                                 ident.1 = "injured",
                                 ident.2 = "uninjured",
                                 logfc.threshold = 0,
                                 min.pct = 0.01)


write.csv(x = LMN_ALL_InjVUninj, 
          file = "LMN_ALL_InjvUninj.csv",
          quote = FALSE)

```

```{r}
saveRDS (LMN_Final, "LMN_Final.rds")
```
